IML Term project paper

### 2022-12-22 paper.rmd

---
title: "IML Term project paper"
author: "Julia Palorinne, Pyry Silomaa, Sara Sippola"
output: pdf_document
urlcolor: blue
date: '`r format(Sys.Date(), "%d.%m.%Y")`'
---

@@

The goal of this assignment is to predict the behavior of a multinomially distributed variable "class4", which describes new particle formation events on specific observation days. The variable is treated as a multinomial variable divided into four separate groups: no new particle formation (referred to in the dataset as "nonevent"), very clear and strong particle formation events (referred to in the dataset as class "Ia"), other particle formation events where the growth and formation rate could be determined with a sufficiently good confidence level (referred to in the dataset as class "Ib") and particle formation events where growth and formation rates could either not be defined or there was a possibility that the assessed rates were not sufficiently accurate. For the most part, we are observing a simpler version of this variable instead in the form of a binomial variable, let's refer to that as "class2", which is simply divided into no new particle formation events happening and new particle formation events happening.

@@

Let's next visualize some our explanatory variables via boxplots and scatterplots to get a deeper understanding of them. For the sake of simplicity, these visualizations grouped by the phenomena being measured. In order to also prelminarily look at how these variables behave in relation to new particle formation events, we'll also look at said measurements behave in relation to variable "class4". In terms of the boxplots, the boxplots are divided into groups by variable class4, and in terms of the scatterplots, the following colors represent different classes: red represents nonevent days, blue represents days with type Ia NPF events, green represents days with type Ib NPF events and black represents days with type II NPF events:

@@

The correlation between the predictors and the value to be predicted is a very natural starting point when it comes to correlation analysis. It is also possible that the predictors themselves are correlated. Many classifiers are affected by high correlation, i.e. *collinearity* between variables. In regression models, for example, it can be difficult to differentiate between the individual effects of collinear variables on the response.

@@

Specifically, the two measures we used to assess the quality of each model were its accuracy and perplexity. We defined accuracy as simply the relative amount of values the model predicted correctly, expressed mathematically for variables of interest $y_i$ and related predictions $\hat{y}_i$ as $\frac{\sum_{i=1}^{n}I(y_i=\hat{y}_i)}{n}$. This measures whether the model defines the labels correctly. All values naturally receieve values within interval $[0,1]$ and higher values imply better results.

@@

However, it is not sufficient to simply define the model on the training dataset and assess accuracy and perplexity on said training dataset as this would woefully inflate the accuracy assessment and deflate the perplexity assessment of the model on a general dataset. We instead attempted to assess accuracy and perplexity through cross-validation methods, specifically leave-one-out cross-validation (loo-cv). The idea behind loo-cv is that for each observation in the training dataset, a model is trained based on all other observations in the dataset. This new model is then used to predict the training dataset value that was left out of defining the model. With this method, we receive based on observations of explanatory variables both predicted values and estimates of probabilities for specific events for each measurement in a way that didn't include the observations in both defining the model training and assessing its quality. Thus, we can use these measures in our accuracy and perplexity calculations to better assess how the model using all the observations in the training dataset behaves on a completely separate test dataset.
```{r preprocess data, echo=FALSE}
# Sample half of the data for training and testing sets
set.seed(42)
idx <- sample(nrow(npfData), nrow(npfData)/2)
npf_train <- npfData[idx, ]
npf_test <- npfData[-idx, ]
```
Next, let's discuss some of the models we considered

@@

Logistic regression models can be used for binomial variables as well as for multinomial variables under specific assumptions. The idea of logistic regression in a binary case is that our variable of interest $Y_i$ are binomially distributed with parameter $q\in [0,1]$ representing probability that $Y_i$ receieves one of the values. Now, with explanatory variables $X_{ij}$ we assume that these $X_{ij}$ influence the parameter $p_i$ for each observation, meaning $p(Y_i|X_i)=q(X_i)^{Y_i}(1-q(X_i))^{1-Y_i}$ is a binomial probability with parameter $q(X_i)$ being of form $\sigma(\sum_j \lambda_j X_{ij})$ for values of individual explanatory variables $X_{ij}$ and coefficients $\lambda_j$ and link function $\sigma: \mathbb{R}\to [0,1]$. Thus we can estimate a logistic regression by finding parameters $\lambda_j$ that maximize the likelihood function $\prod p(Y_i|X_i)$.

Notable benefits of the logistic regression model include that as a discriminative model it eventually reaches a lower asymptotic variance than comparable generative models. On top of this, compared to for example comparable support vector classifiers, on average logistic regression models fare better under situations where the classes are not easily separable by explanatory variables.

@@

Variance inflation factors (VIFs) can be calculated for each explanatory variable in a model through assessing linear models for each of the explanatory variables. The VIF for explanatory variable $X_j$ is defined as $\frac{1}{1-R_i^2}$ where $R_i^2$ is the amount of variance of $X_j$ explained through a linear model by the other explanatory variables. Thus, these values express how much the explanatory variable in question can be described by a linear combination of the other explanatory variables. Thus, particularly high VIFs imply multicollinearity being present in the model. We calculated VIFs for our model with function vif from R package car.

@@

Lasso regression has obvious links to logistic regression but it includes a penalty term which punishes models with high values of coefficients. In comparable linear regression minimizing squared error, the minimizable function is the sum of the squared error terms and $\sum_{i=1}^n \gamma |\lambda_i|$ where $\gamma>0$. We similarly as in logistic regression assess coefficients for a model that describes parameter $q$ of a binomial model that minimizes this function.

One of the benefits of lasso regression is that including this penalty term in the optimization process leads to simpler models with fewer explanatory variables, as also lower-dimensional models lead are preferred. Thus, the lasso model in and of itself can remove variables which are highly correlated with other variables from the model and thus at least alleviate multicollinarity within the model. However, according to the High-Dimensional Statistics course material, selection between possible correlated variables might with extremely highly correlated variables break down due to there being multiple optima. Thus, it seems reasonable to note that with as highly correlated variables as the ones we're dealing with, the model might not properly describe the data-generating process but instead simply choose the variables that happen to create a slightly fit better on the training data.

In terms of our modelling, we decided to use lasso instead of a comparable method called ridge regression, which uses penalty term $\sum_{j=1}^m \gamma \lambda_j^2$. The reason for this is that we preferred to make a simpler model with fewer explanatory variables rather than ridge regression, which tends to not reduce the dimension of the model but rather simply give similar coefficients to all correlated explanatory variables. In order to estimate the model, we used R functions glmnet and cv.glmnet from package glmnet. We attempted to assess the value of $\gamma$ which would minimize loo cross-validated misclassification rate with the function cv.glmnet. After defining this value, we calculated cross-validated accuracy (which should be equivalent to what was calculated by cv.glmnet) as well as the cross-validated perplexity for the model estimated by glmnet.

@@

The idea of the binary support vector classifier is based on the notion of finding within a $p$-dimensional problem a $(p-1)$-dimensional hyperplane $\lambda_0+\sum\lambda_i x_i=0$ to divide the set of explanatory variables. Now, for this hyperplane, we can divide our observations into two groups, those values for which $\lambda_0+\sum_{i=1}^m\lambda_j x_i>0$ and those values for which $\lambda_0+\sum\lambda_i x_i<0$. Thus, this can function as a model through which we can predict binary variables such as our variable of interest.

However, defining such hyperplanes can be slightly difficult as all values might not be separable by a hyperplane or the hyperplane that separates values can overfit to training data. Thus, we define a soft-margin classifier, which has the benefits of letting some values be misclassified or be closer to the hyperplane than a strict maximal margin classifier would define, which leads to a more robust model that fits most of the training dataset better.

In order to define this, we need to find the optimal hyperplane to separate the observations. This optimization problem is an optimization problem of $M$ in relation to coefficients $\lambda_i$ and slack variables $\epsilon$ within inequation $y_i(\lambda_0+\sum_{j=1}^m\lambda_jx_{ij})\ge M(1-\epsilon_i)$ with limitations $\sum_{j=1}^{m}\lambda_j^2=1$, $\epsilon_i\ge0$ and $\sum_{i=1}^n\epsilon_i\le C$ for tuning parameter $C$.

As in our optimization the data only depends on the inner products of the explanatory variables and the values of the variable of interest, we can use different generalizations of inner products to gain different kinds of models. Thus with different choices of kernels we can model different kinds of data.

Other benefits benefits of using support vector classifiers include that they have been proven to behave well in several different applications and that it has been shown that on average support vector classifiers perform better than logistic regression models in problems where the classes are well-separated. However, the model can vary quite notably based on the tuning parameter $C$ as this parameter affects the amount of observations that violate the margin and thus increases or decreases the amount of support vectors that influence the model.

For our model, we used functions tune and svm from R package e1071. We chose to run a 10-fold cross validation on a set of tuning parameters $C$ to assess which kernel performed best and chose to model the value based on that. In these preliminary assessments, we assessed the radial and linear kernels had the best performance of the ones available in package e1071. For these kernels, we ran loo-cross validation to assess the value of tuning parameter $C$ and calculated for said tuning parameter loo-cross-validated accuracy and perplexity. Eventually, the radial kernel had slightly better accuracy but the linear kernel had lower perplexity and as a result, we chose to use the linear kernel in our model.

@@

In order to reduce knitting time for the file, we will in this preliminary report instead of expressing the full R code to calculate the values, simply report the results we received and instead return the full R code only with the final report. The results for the two-class models in terms of loo-cross-validated accuracy and perplexity are:
\begin{table}
\begin{tabular}{|r|cc|}
\hline
Model type & loo-cv accuracy & loo-cv perplexity\\
\hline
Dummy model & 0 & 2.01\\
Logistic regression & 0.89 & 1.30\\
Naive Bayes & 0.81 & 152.85\\
LDA & 0.89 & 1.33\\
QDA & 0.85 & 6.34\\
SVM & 0.88 & 1.34\\
Lasso & 0.89 & 1.29\\
Random Forest & 0.89 & 1.35\\
\hline
\end{tabular}
\end{table}

@@

With our observed best-performing models being the logistic regression model, the random forest model, the Lasso regression model and the linear discriminant analysis. Out of these models, we eventually chose to use the logistic regression model. The Lasso model was our best-performing model, but we eventually chose not to use it for practical reasons related to computational heaviness in calculating the multi-class model. Mainly, the amount of information we could gain through calculations in R without R crashing was relatively limited, to the degree that we couldn't always even calculate the cross-validated best choice for the tuning parameter.

@@

Thus we were left with the logistic regression model, which was also lower-dimensional than many of the other models with only 10 explanatory variables. However, we could question in relation to this model whether there is sufficient basis to for example remove certain variables, which might describe relevant information, as well as for example whether including an intercept in the model would be meaningful in terms of interpretation. Currently, several of these decisions are based on likelihood-ratio tests performed within the R function glm, which was pursued in an attempt to create a lower-dimensional model. Also, comparing the logistic regression model to for example the Lasso, which does variable selection on a model with a somewhat comparable minimzable loss function.

## Multiclass-classifier
we will extend it into a multinomial logistic regression model in order to assess the multi-class accuracy for our model. The basic idea of multinomial logistic regression here is that out of $K$ classes we choose a baseline value of our variable of interest, and compare other values to it, leading to $p(Y=k|X=x)=\frac{e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}{1+\sum_{k=1}^{K-1} e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}$ where our baseline class is class $K$. Alternatively, we can use equivalent so-called softmax coding where we  treat all classes equivalently and thus $p(Y=k|X=x)=\frac{e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}{\sum_{k=1}^K-1 e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}$. The interpretation of logistic regression coefficients comes now from the following: $\log(\frac{p(\hat{y}_i=k_a)}{p(\hat{y_i}=k_b)})=(\lambda_{a0}-\lambda_{b0})+\sum_{j=1}^n(\lambda_{aj}-\lambda_{bj})x_{ij}$. Thus the interpretation of these coefficients is in relation to the difference of the coefficients for different classes and describe the ratio between the probabilities that our variable of interest receives a value in one of these two classes.

@@

In our model, we used the same dataset as in our binary logistic regression model, in order to avoid issues of multicollinearity. In order to define this model, we used function multinom from R package nnet with softmax codin. This creates estimated probabilities $\hat{p}(\hat{y}_i=y)$ for all four classes $y$, and through that we could calculate loo-cross-validated values for accuracy as well as perplexity.

>>>>

---
title: "IML Term project paper"
author: "Julia Palorinne, Pyry Silomaa, Sara Sippola"
output: pdf_document
bibliography: IML.bib
csl: apa.csl
urlcolor: blue
date: '`r format(Sys.Date(), "%d.%m.%Y")`'
---

@@

The goal of this assignment is to predict the behavior of a multinomially distributed variable "class4", which describes new particle formation events on specific observation days. The variable is treated as a multinomial variable divided into four separate groups which are according to the article "Formation and growth of fresh atmospheric aerosols: eight years of aerosol size distribution data from SMEAR II, Hyytiälä, Finland" [@eventstat]: no new particle formation (referred to in the dataset as "nonevent"), very clear and strong particle formation events (referred to in the dataset as class "Ia"), other particle formation events where the growth and formation rate could be determined with a sufficiently good confidence level (referred to in the dataset as class "Ib") and particle formation events where growth and formation rates could either not be defined or there was a possibility that the assessed rates were not sufficiently accurate. For the most part, we are observing a simpler version of this variable instead in the form of a binomial variable, let's refer to that as "class2", which is simply divided into no new particle formation events happening and new particle formation events happening.

@@

Let's next visualize with R [@RDCT2008] some our explanatory variables via boxplots and scatterplots to get a deeper understanding of them. For the sake of simplicity, these visualizations grouped by the phenomena being measured. In order to also prelminarily look at how these variables behave in relation to new particle formation events, we'll also look at said measurements behave in relation to variable "class4". In terms of the boxplots, the boxplots are divided into groups by variable class4, and in terms of the scatterplots, the following colors represent different classes: red represents nonevent days, blue represents days with type Ia NPF events, green represents days with type Ib NPF events and black represents days with type II NPF events:

@@

The correlation between the predictors and the value to be predicted is a very natural starting point when it comes to correlation analysis. It is also possible that the predictors themselves are correlated. Many classifiers are affected by high correlation, i.e. *collinearity* between variables. In regression models, for example, it can be difficult to differentiate between the individual effects of collinear variables on the response. Let's visualize these correlations with R package [@corrplot]

@@

Specifically, the two measures we used to assess the quality of each model were its accuracy and perplexity. We defined accuracy similarly as in the course material [@book, p. 37] as simply the relative amount of values the model predicted correctly, expressed mathematically for variables of interest $y_i$ and related predictions $\hat{y}_i$ as $\frac{\sum_{i=1}^{n}I(y_i=\hat{y}_i)}{n}$. This measures whether the model defines the labels correctly. All values naturally receieve values within interval $[0,1]$ and higher values imply better results.

@@

However, it is not sufficient to simply define the model on the training dataset and assess accuracy and perplexity on said training dataset as this would woefully inflate the accuracy assessment and deflate the perplexity assessment of the model on a general dataset. We instead attempted to assess accuracy and perplexity through cross-validation methods, specifically leave-one-out cross-validation (loo-cv). The idea behind loo-cv according to the course material [@book, p. 200-202] is that for each observation in the training dataset, a model is trained based on all other observations in the dataset. This new model is then used to predict the training dataset value that was left out of defining the model. With this method, we receive based on observations of explanatory variables both predicted values and estimates of probabilities for specific events for each measurement in a way that didn't include the observations in both defining the model training and assessing its quality. Thus, we can use these measures in our accuracy and perplexity calculations to better assess how the model using all the observations in the training dataset behaves on a completely separate test dataset.
```{r preprocess data, echo=FALSE}
# Sample half of the data for training and testing sets
set.seed(42)
idx <- sample(nrow(npfData), nrow(npfData)/2)
npf_train <- npfData[idx, ]
npf_test <- npfData[-idx, ]
```
Next, let's discuss some of the models we considered

@@

According to the course material [@book, p. 133-137], Logistic regression models can be used for binomial variables as well as for multinomial variables under specific assumptions. The idea of logistic regression in a binary case is that our variable of interest $Y_i$ are binomially distributed with parameter $q\in [0,1]$ representing probability that $Y_i$ receieves one of the values. Now, with explanatory variables $X_{ij}$ we assume that these $X_{ij}$ influence the parameter $p_i$ for each observation, meaning $p(Y_i|X_i)=q(X_i)^{Y_i}(1-q(X_i))^{1-Y_i}$ is a binomial probability with parameter $q(X_i)$ being of form $\sigma(\sum_j \lambda_j X_{ij})$ for values of individual explanatory variables $X_{ij}$ and coefficients $\lambda_j$ and link function $\sigma: \mathbb{R}\to [0,1]$. Thus we can estimate a logistic regression by finding parameters $\lambda_j$ that maximize the likelihood function $\prod p(Y_i|X_i)$.

Notable benefits of the logistic regression model include that according to the article "On Discriminative vs. Generative classifiers: A comparison of logistic regression and naive Bayes"[@ng2001] as a discriminative model it eventually reaches a lower asymptotic variance than comparable generative models. On top of this, according to the course material [book, p. 387], compared to for example comparable support vector classifiers, on average logistic regression models fare better under situations where the classes are not easily separable by explanatory variables.

@@

According to the course material [@book, p. 102], Variance inflation factors (VIFs) can be calculated for each explanatory variable in a model through assessing linear models for each of the explanatory variables. The VIF for explanatory variable $X_j$ is defined as $\frac{1}{1-R_i^2}$ where $R_i^2$ is the amount of variance of $X_j$ explained through a linear model by the other explanatory variables. Thus, these values express how much the explanatory variable in question can be described by a linear combination of the other explanatory variables. Thus, particularly high VIFs imply multicollinearity being present in the model. We calculated VIFs for our model with function vif from R package car.

@@

Lasso regression has obvious links to logistic regression but it includes a penalty term which punishes models with high values of coefficients. According to the course material [@book, p. 241-242] in comparable linear regression minimizing squared error, the minimizable function is the sum of the squared error terms and $\sum_{i=1}^n \gamma |\lambda_i|$ where $\gamma>0$. We similarly as in logistic regression assess coefficients for a model that describes parameter $q$ of a binomial model that minimizes this function.

One of the benefits of Lasso regression according to the High-Dimensional Statistics course material [@Lasso] is that including this penalty term in the optimization process with correlated variables leads to simpler models with fewer explanatory variables, as also lower-dimensional models lead are preferred. Thus, the lasso model in and of itself can remove variables which are highly correlated with other variables from the model and thus at least alleviate multicollinarity within the model. However, according to the High-Dimensional Statistics course material [@Lasso], selection between possible correlated variables might with extremely highly correlated variables break down due to there being multiple optima. Thus, it seems reasonable to note that with as highly correlated variables as the ones we're dealing with, the model might not properly describe the data-generating process but instead simply choose the variables that happen to create a slightly fit better on the training data.

In terms of our modelling, we decided to use Lasso instead of a comparable method called ridge regression, which according to the course material [@book, p. 237-238] uses penalty term $\sum_{j=1}^m \gamma \lambda_j^2$. The reason for this is that we preferred to make a simpler model with fewer explanatory variables rather than ridge regression, which tends to not reduce the dimension of the model but rather simply give similar coefficients to all correlated explanatory variables. In order to estimate the model, we used R functions glmnet and cv.glmnet from package glmnet [@glmnet]. We attempted to assess the value of $\gamma$ which would minimize loo cross-validated misclassification rate with the function cv.glmnet. After defining this value, we calculated cross-validated accuracy (which should be equivalent to what was calculated by cv.glmnet) as well as the cross-validated perplexity for the model estimated by glmnet.

@@

The idea of the binary support vector classifier is according to the course material [@book, p. 369-370] based on the notion of finding within a $p$-dimensional problem a $(p-1)$-dimensional hyperplane $\lambda_0+\sum\lambda_i x_i=0$ to divide the set of explanatory variables. Now, for this hyperplane, we can divide our observations into two groups, those values for which $\lambda_0+\sum_{i=1}^m\lambda_j x_i>0$ and those values for which $\lambda_0+\sum\lambda_i x_i<0$. Thus, this can function as a model through which we can predict binary variables such as our variable of interest.

However, defining such hyperplanes can be slightly difficult as all values might not be separable by a hyperplane or the hyperplane that separates values can overfit to training data. Thus, according to the course material [@book, p. 373-375], we can define a soft-margin classifier, which has the benefits of letting some values be misclassified or be closer to the hyperplane than a strict maximal margin classifier would define, which leads to a more robust model that fits most of the training dataset better.

In order to define this, we need to find the optimal hyperplane to separate the observations. According to the course material [@book, p. 375-378], this optimization problem is an optimization problem of $M$ in relation to coefficients $\lambda_i$ and slack variables $\epsilon$ within inequation $y_i(\lambda_0+\sum_{j=1}^m\lambda_jx_{ij})\ge M(1-\epsilon_i)$ with limitations $\sum_{j=1}^{m}\lambda_j^2=1$, $\epsilon_i\ge0$ and $\sum_{i=1}^n\epsilon_i\le C$ for tuning parameter $C$.

According to the course material [@book, p. 380-383], our optimization the data only depends on the inner products of the explanatory variables and the values of the variable of interest. Thus we can use different generalizations of inner products to gain different kinds of models. Because of this, with different choices of kernels we can model different kinds of data.

Other benefits benefits of using support vector classifiers according to the course material [@book, p. 367, 387] include that they have been proven to behave well in several different applications and that it has been shown that on average support vector classifiers perform better than logistic regression models in problems where the classes are well-separated. However, the model can vary quite notably based on the tuning parameter $C$ as this parameter affects the amount of observations that violate the margin and thus increases or decreases the amount of support vectors that influence the model.

For our model, we used functions tune and svm from R package e1071 [@e1071]. We chose to run a 10-fold cross validation on a set of tuning parameters $C$ to assess which kernel performed best and chose to model the value based on that. In these preliminary assessments, we assessed the radial and linear kernels had the best performance of the ones available in package e1071. For these kernels, we ran loo-cross validation to assess the value of tuning parameter $C$ and calculated for said tuning parameter loo-cross-validated accuracy and perplexity. Eventually, the radial kernel had slightly better accuracy but the linear kernel had lower perplexity and as a result, we chose to use the linear kernel in our model.

@@

In order to reduce knitting time for the file, we will in this report instead of expressing the full R code to calculate the values, simply report the results we received. The results for the two-class models in terms of loo-cross-validated accuracy and perplexity are:
\begin{table}
\begin{tabular}{|r|cc|}
\hline
Model type & loo-cv accuracy & loo-cv perplexity\\
\hline
Dummy model & 0 & 2.01\\
Logistic regression & 0.89 & 1.30\\
Naive Bayes & 0.81 & 152.85\\
LDA & 0.89 & 1.33\\
QDA & 0.85 & 6.34\\
SVM & 0.88 & 1.34\\
Lasso & 0.89 & 1.29\\
Random Forest & 0.89 & 1.35\\
\hline
\end{tabular}
\end{table}

@@

With our observed best-performing models being the logistic regression model, the random forest model, the Lasso regression model and the linear discriminant analysis. Out of these models, we eventually chose to use the logistic regression model. The Lasso model was our best-performing model, but we eventually chose not to use it for practical reasons related to computational heaviness in calculating the multi-class model. Mainly, the amount of information we could gain through calculations in R without R crashing was relatively limited, to the degree that we couldn't always even calculate the cross-validated best choice for the tuning parameter. We could consider running the models again with a smaller amount of folds in the cross-validation, as that might according to the course material [@book] give more accurate estimates, and thus the model would be runnable. However, for the moment we preferred our current approach as with Lasso we couldn't choose which variables to use in the model, which might not be preferrable in further analyses.

@@

Thus we were left with the logistic regression model, which was also lower-dimensional than many of the other models with only 10 explanatory variables. However, we could question in relation to this model whether there is sufficient basis to for example remove certain variables, which might describe relevant information, as well as for example whether including an intercept in the model would be meaningful in terms of interpretation. Currently, several of these decisions are based on Wald tests performed within the R [@RDCT2008] function glm, which was pursued in an attempt to create a lower-dimensional model.

## Multiclass-classifier
We will extend the logistic regression model into a multinomial logistic regression model in order to assess the multi-class accuracy for our model. According to the course material [@book, p. 140-141], the basic idea of multinomial logistic regression here is that out of $K$ classes we choose a baseline value of our variable of interest, and compare other values to it, leading to $p(Y=k|X=x)=\frac{e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}{1+\sum_{k=1}^{K-1} e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}$ where our baseline class is class $K$. Alternatively, we can use equivalent so-called softmax coding where we treat all classes equivalently and thus $p(Y=k|X=x)=\frac{e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}{\sum_{k=1}^K-1 e^{\beta_{k0}+\sum_{j=1}^m\beta_{kj}x_{ij}}}$. The interpretation of logistic regression coefficients comes now from the following: $\log(\frac{p(\hat{y}_i=k_a)}{p(\hat{y_i}=k_b)})=(\lambda_{a0}-\lambda_{b0})+\sum_{j=1}^n(\lambda_{aj}-\lambda_{bj})x_{ij}$. Thus the interpretation of these coefficients is in relation to the difference of the coefficients for different classes and describe the ratio between the probabilities that our variable of interest receives a value in one of these two classes.

@@

In our model, we used the same dataset as in our binary logistic regression model, in order to avoid issues of multicollinearity. In order to define this model, we used function multinom from R package nnet [@nnet] with softmax coding. This creates estimated probabilities $\hat{p}(\hat{y}_i=y)$ for all four classes $y$, and through that we could calculate loo-cross-validated values for accuracy as well as perplexity.
